---
title: "ES 207 HW2"
author: "Erik Bolch"
output: html_document
---
<span style ="color:crimson">1. Using the help documentation, provide an alternative useage for the function ncol()</span>

<span style="color:blue"> An alternative use of ncol() could be to determine if an object has dimensions associated with it. For example when ncol function is used on a vector as opposed to an array or matrix it will return NULL.</span>

```{r}
?ncol
m<-matrix(1:10, 2,5)
m
ncol(m)
NCOL(m)

v<-1:10
v
ncol(v)
NCOL(v)
```
2. Make sure to clone the tutorial repo to a new repo in your GitHub account called es207_hw2. I will check your repo, so don’t skip this step!

3. Convert RMarkdown_Demo_1.R into a well commented and easy to follow R Markdown document. Turn this in as a .pdf document separate from the rest of your homework assignment. Call it “yourlastname_demo_1.pdf”.

4. Complete the remaining homework below as a well-commented and easy to follow R Markdown notebook with approriate accompanying text. Turn this in as a .html document. Call it “yourlastname_hw2.html”.


```{r}

```

```{r}

```


