---
title: "RMarkdown_Demo_1"
author: "Erik Bolch"
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
# Example R Markdown Script
### Adapted from: Tidy data and efficient manipulation coding club tutorial
##### January 18th 2017
##### Sandra Angers-Blondin (s.angers-blondin@ed.ac.uk)
##### John Godlee
##### 24/Jan/2017

Install and load the relevant packages
<p style="color:red">Changed tidyr to tidyverse</p>
```{r}
library(dplyr) # an excellent data manipulation package
library(tidyverse) # a package to format your data
library(pander) #to create pretty tables
```
Set your working directory to the folder where you have downloaded the datasets
<p style="color:red">Commented out because not recommended for Rmd files</p>
```{r}
#setwd("./data/CC-3-DataManip-master")

```
Import data
<p style="color:red">Germination file didn't exist, changed to dragons which is included.</p>
```{r}

elongation <- read_csv("./data/CC-3-DataManip-master/EmpetrumElongation.csv") # stem elongation measurements on Empetrum
dragons <- read_csv("./data/CC-3-DataManip-master/dragons.csv") # dragons subjected to spices

```
Tidying the data and Putting the data into long format using gather()

```{r}
elongation_long <- gather(elongation, Year, Length, c(X2007, X2008, X2009, X2010, X2011, X2012)) 
  #gather() works like this: data, key, value, columns to gather. Here we want the lengths (value) to be gathered by year (key). Note that you are completely making up the names of the second and third arguments, unlike most functions in R.
head(elongation_long)
```
Investigating the data
Create a boxplot of `elongation_long' to visualise elongation for each year.

This set of boxplots can be added to your R Markdown document by putting the code in a code chunk
```{r}
boxplot(Length ~ Year, 
        data = elongation_long, 
        xlab = "Year", 
        ylab = "Elongation (cm)", 
        main = "Annual growth of Empetrum hermaphroditum")
```



Use filter() to keep only the rows of 'dragon' for species 'hungarian_horntail' 
```{r}
dragon.hh <- filter(dragons, species == 'hungarian_horntail')
```



Let's have a look at the distribution of dragons species exposed to spices
This histogram can be added to your R Markdown document by simply putting the code in a code chunk
Try adding some plain text to your R markdown document to explain the histogram
```{r}
hist(dragon.hh$wasabi, breaks = 8)
```

Use mutate() to create a new column of the dragon percentage using the total number total and the number of dragons exposed to jalapeno out of total spice exposure

```{r}
dragon_j_percent <- mutate(dragons, Percent_Jalapeno = jalapeno / (tabasco+jalapeno+wasabi+paprika) * 100)
head(dragon_j_percent)
```

Use a pipe to get a table of summary statistics for each Seed type

```{r}
dragon_summ <- dragon_j_percent %>%
  group_by(species) %>%
  summarise("Mean jalapeno exposure per" = mean(jalapeno), "Max jalapeno exposure per" = max(jalapeno), "Min jalapeno exposure" = min(jalapeno))
```

Make a table of 'dragon_summ' in your R markdown document using pander(), the instructions can be found in the tutorial

```{r}
pander(dragon_summ)
```























